---
title: "TRI Google Ad Grants Dashboard (Static)"
subtitle: "CRM + TRM search demand, match-type mix, and query themes"
author: "Angela Qian"
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    toc-expand: true
    embed-resources: true
    code-fold: true
    code-tools: true
    dashboard:
      orientation: rows
      nav-buttons: true
execute:
  echo: true
  warning: false
  message: false
---

```{r}
library(tidyverse)
library(janitor)
library(scales)
library(stringr)
```

```{r}
campaign_files <- c(
  CRM = "/Users/angelaqian/Campaign report (CRM).csv",
  TRM = "/Users/angelaqian/Campaign report (TRM).csv"
)

terms_files <- c(
  CRM = "/Users/angelaqian/Search terms report (CRM).csv",
  TRM = "/Users/angelaqian/Search terms report (TRM).csv"
)

keyword_files <- c(
  CRM = "/Users/angelaqian/Search keyword report (CRM).csv",
  TRM = "/Users/angelaqian/Search keyword report (TRM).csv"
)
```
```{r}
read_ads_csv_skip2 <- function(path) {
  if (!file.exists(path)) stop("File not found: ", path)

  txt <- readLines(path, encoding = "UTF-16LE", warn = FALSE)
  txt <- txt[!is.na(txt)]

  tmp <- tempfile(fileext = ".csv")
  writeLines(txt, tmp, useBytes = TRUE)

  df <- readr::read_csv(
    tmp,
    skip = 2,
    show_col_types = FALSE,
    progress = FALSE,
    guess_max = 50000
  ) %>% clean_names()

  if (ncol(df) == 1) {
    stop(
      "Imported as ONE column (bad delimiter/export). Re-export as CSV (comma).\n",
      "Single column name: ", names(df)[1]
    )
  }
  df
}

to_num <- function(x) {
  x <- as.character(x)
  x <- gsub("%", "", x)
  x <- gsub("\\$", "", x)
  x <- gsub(",", "", x)
  suppressWarnings(as.numeric(x))
}

drop_totals <- function(df) {
  nm <- names(df)
  label_col <- intersect(c("campaign","search_term","ad_group","campaign_name","search_terms_report"), nm)
  if (length(label_col) == 0) return(df)
  col <- label_col[1]
  df %>% filter(!str_detect(tolower(.data[[col]]), "^total:"))
}

pick_col <- function(df, candidates) {
  hit <- intersect(candidates, names(df))
  if (length(hit) == 0) return(NA_character_)
  hit[1]
}

standardize_campaign <- function(df) {
  col_campaign <- pick_col(df, c("campaign","campaign_name","campaign_report"))
  col_impr     <- pick_col(df, c("impr","impressions"))
  col_clicks   <- pick_col(df, c("clicks"))
  col_cost     <- pick_col(df, c("cost"))
  col_conv     <- pick_col(df, c("conversions","conv","all_conversions"))
  col_ctr      <- pick_col(df, c("ctr"))
  col_cpc      <- pick_col(df, c("avg_cpc","average_cpc"))
  col_cvr      <- pick_col(df, c("conv_rate","conversion_rate"))
  col_cpa      <- pick_col(df, c("cost_conv","cost_per_conv","cost_per_conversion"))

  if (is.na(col_campaign)) stop("Could not find campaign column. Columns: ", paste(names(df), collapse=", "))

  df %>%
    drop_totals() %>%
    mutate(
      campaign    = as.character(.data[[col_campaign]]),
      clicks      = if (!is.na(col_clicks)) to_num(.data[[col_clicks]]) else NA_real_,
      cost        = if (!is.na(col_cost))   to_num(.data[[col_cost]])   else NA_real_,
      impressions = if (!is.na(col_impr))   to_num(.data[[col_impr]])   else NA_real_,
      conversions = if (!is.na(col_conv))   to_num(.data[[col_conv]])   else 0,
      ctr         = if (!is.na(col_ctr))    to_num(.data[[col_ctr]])    else NA_real_,
      avg_cpc     = if (!is.na(col_cpc))    to_num(.data[[col_cpc]])    else NA_real_,
      conv_rate   = if (!is.na(col_cvr))    to_num(.data[[col_cvr]])    else NA_real_,
      cost_conv   = if (!is.na(col_cpa))    to_num(.data[[col_cpa]])    else NA_real_
    ) %>%
    filter(!is.na(campaign), campaign != "", campaign != "--")
}

standardize_terms <- function(df) {
  col_term     <- pick_col(df, c("search_term","search_terms","term","query","search_terms_report"))
  col_match    <- pick_col(df, c("match_type","keyword_match_type"))
  col_campaign <- pick_col(df, c("campaign","campaign_name"))
  col_impr     <- pick_col(df, c("impr","impressions"))
  col_clicks   <- pick_col(df, c("clicks"))
  col_cost     <- pick_col(df, c("cost"))
  col_conv     <- pick_col(df, c("conversions","conv","all_conversions"))
  col_ctr      <- pick_col(df, c("ctr"))

  if (is.na(col_term)) stop("Could not find search term column. Columns: ", paste(names(df), collapse=", "))

  df %>%
    drop_totals() %>%
    mutate(
      search_term = as.character(.data[[col_term]]),
      match_type  = if (!is.na(col_match)) as.character(.data[[col_match]]) else NA_character_,
      campaign    = if (!is.na(col_campaign)) as.character(.data[[col_campaign]]) else NA_character_,
      clicks      = if (!is.na(col_clicks)) to_num(.data[[col_clicks]]) else NA_real_,
      cost        = if (!is.na(col_cost))   to_num(.data[[col_cost]])   else NA_real_,
      impressions = if (!is.na(col_impr))   to_num(.data[[col_impr]])   else NA_real_,
      conversions = if (!is.na(col_conv))   to_num(.data[[col_conv]])   else 0,
      ctr         = if (!is.na(col_ctr))    to_num(.data[[col_ctr]])    else NA_real_
    ) %>%
    filter(!is.na(search_term), search_term != "", search_term != "--")
}
```
```{r}
campaign <- imap_dfr(campaign_files, ~ read_ads_csv_skip2(.x) %>% standardize_campaign() %>% mutate(program = .y))
terms    <- imap_dfr(terms_files,    ~ read_ads_csv_skip2(.x) %>% standardize_terms()    %>% mutate(program = .y))
keywords <- imap_dfr(keyword_files,  ~ read_ads_csv_skip2(.x) %>% clean_names()          %>% mutate(program = .y))
```

## Overview

::: card
What this dashboard shows

This dashboard summarizes TRI’s Google Ad Grants activity for CRM vs TRM using the latest exports in /data/. It focuses on volume, match-type mix, and top search themes (for non-brand demand + negative keyword opportunities).
:::

```{r}
kpi <- campaign %>%
  group_by(program) %>%
  summarise(
    clicks = sum(clicks, na.rm=TRUE),
    cost = sum(cost, na.rm=TRUE),
    avg_cpc = cost / clicks,
    conv = sum(conversions, na.rm=TRUE),
    .groups = "drop"
  )

knitr::kable(
  kpi %>% mutate(cost = dollar(cost), avg_cpc = dollar(avg_cpc), clicks = comma(clicks), conv = comma(conv)),
  align = "l"
)
```

## Campaign Performance (CRM vs TRM)

### Clicks by campaign

```{r}
p_clicks <- campaign %>%
  group_by(program, campaign) %>%
  summarise(clicks = sum(clicks, na.rm=TRUE), .groups="drop") %>%
  arrange(program, clicks) %>%
  ggplot(aes(x = clicks, y = fct_reorder(campaign, clicks))) +
  geom_col() +
  facet_wrap(~program, scales="free_y") +
  scale_x_continuous(labels = comma) +
  labs(title = "Clicks by Campaign", x = "Clicks", y = NULL)
p_clicks
```

### Spend by campaign

```{r}
p_cost <- campaign %>%
  group_by(program, campaign) %>%
  summarise(cost = sum(cost, na.rm=TRUE), .groups="drop") %>%
  arrange(program, cost) %>%
  ggplot(aes(x = cost, y = fct_reorder(campaign, cost))) +
  geom_col() +
  facet_wrap(~program, scales="free_y") +
  scale_x_continuous(labels = dollar) +
  labs(title = "Spend by Campaign", x = "Cost ($)", y = NULL)
p_cost
```

## Match-type mix (quality proxy)

### Click share by match type (CRM vs TRM)

```{r}
match_share_prog <- terms %>%
  filter(!is.na(match_type), match_type != "") %>%
  group_by(program, match_type) %>%
  summarise(clicks = sum(clicks, na.rm=TRUE), .groups="drop") %>%
  group_by(program) %>%
  mutate(share = clicks / sum(clicks)) %>%
  ungroup()

p_match_prog <- ggplot(match_share_prog, aes(x = share, y = fct_reorder(match_type, share))) +
  geom_col() +
  facet_wrap(~program) +
  geom_text(aes(label = percent(share, accuracy = 0.1)), hjust = -0.1, size=3.5) +
  scale_x_continuous(labels = percent, expand = expansion(mult = c(0, 0.2))) +
  labs(title = "Match-Type Mix by Program", subtitle = "Broad vs Phrase vs Exact", x = "Share of Clicks", y = NULL)

p_match_prog
```

## Search Demand Themes (top queries)

### Top search terms by program

```{r}
top_terms_prog <- terms %>%
  group_by(program, search_term) %>%
  summarise(clicks = sum(clicks, na.rm=TRUE), .groups="drop") %>%
  arrange(program, desc(clicks)) %>%
  group_by(program) %>%
  slice_head(n = 8) %>%
  ungroup() %>%
  mutate(search_term = str_wrap(search_term, 28))

p_terms_prog <- ggplot(top_terms_prog, aes(x = clicks, y = fct_reorder(search_term, clicks))) +
  geom_col() +
  facet_wrap(~program, scales="free_y") +
  geom_text(aes(label = comma(clicks)), hjust = -0.1, size=3.3) +
  scale_x_continuous(labels = comma, expand = expansion(mult = c(0, 0.2))) +
  labs(title = "Top Search Terms by Program", subtitle = "Good for non-brand themes + negatives", x = "Clicks", y = NULL)

p_terms_prog
```

::: callout-note
Interpretation notes

-   Broad match can uncover new demand, but usually needs strong negative keywords to avoid low-intent clicks.
-   If conversions show as 0 in export, treat this as a traffic-mix scan (not ROI). :::
:::
